Manuscript

### 2022-09-01 analysis/manuscript.Rmd
change C plot to be sum of C not max-C

tar_load(Neale_SGG_dir_filt)
tar_load(traits_corr_spearman)
tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders_pos)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)
#tar_load(potential_trait_confounders_MVMR)
tar_load(corr_mat_traits)
tar_load(corr_impact_by_traits_causal)
tar_load(household_MR_SNPmeta_sensitivity)
tar_load(binned_pheno_corrs)

@@

sig_confounders <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding) %>%
  mutate(total_potential_confounders = n()) %>%
  dplyr::summarize(description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  dplyr::rename(trait_description = description) %>% 
  dplyr::select(trait_description, trait_ID, everything())

sig_confounders_top5 <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 5) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(trait_description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   top5_potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  left_join(corr_vs_MR_effects_sig_dat.DT, by = c("trait_description", "trait_ID")) %>%
  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval,everything())

@@

@@

max_C_data <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 1) %>%
  left_join(couple_MR_vs_trait_corr %>% 
              dplyr::rename(outcome_ID_IVW_beta = IVW_beta) %>%
              dplyr::rename(outcome_ID_IVW_se = IVW_se) %>%
              dplyr::rename(outcome_ID_couple_r = couple_r) %>% 
              dplyr::rename(outcome_ID_couple_r_se = couple_r_se) %>% 
              dplyr::select(outcome_ID, outcome_ID_IVW_beta, outcome_ID_IVW_se, outcome_ID_couple_r, outcome_ID_couple_r_se)) %>%
  mutate(same_trait_corr_minus_MR_est = abs(outcome_ID_couple_r)- abs(outcome_ID_IVW_beta)) %>%

  dplyr::rename(Z_beta = IVW_beta) %>%
  dplyr::rename(Z_se = IVW_se) %>%
  dplyr::rename(Z_pval = IVW_pval) %>%
  mutate(Z_sq_beta = Z_beta^2) %>%
  mutate(Z_sq_se = sqrt(4*Z_beta^2*Z_se^2+2*Z_se^4)) %>%
  dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
  dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
  mutate(Z_prod = Z_sq_beta*Z_AM_beta) %>%
  mutate(Z_prod_var = variance_of_product(Z_sq_beta, Z_sq_se, Z_AM_beta, Z_AM_se)) %>%
  mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
  mutate(same_trait_corr_minus_MR_est_se = sqrt(variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))) %>%
  mutate(Phenotype = outcome_description) %>% 
  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
                               Phenotype == "Pulse rate, automated reading" ~ "Pulse rate",
                               Phenotype == "Body fat percentage" ~ "BFP",
                               Phenotype == "Standing height" ~ "Height",
                               Phenotype == "Reticulocyte percentage" ~ "Reticulocyte %",

@@

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - 1.96*same_trait_corr_minus_MR_est_se, 
                 ymax = same_trait_corr_minus_MR_est + 1.96*same_trait_corr_minus_MR_est_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = corr_due_to_confounding - 1.96*Z_prod_se, 
                 xmax = corr_due_to_confounding + 1.96*Z_prod_se),
    colour = "grey",
    width = 0
  ) +

@@

@@

## Supplementary tables
write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Table 1 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Table 2 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Table 3 pheno_binned_corrs", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Table 4 sig_corr_vs_MR", row.names=FALSE)
write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Table 5 sig_confounders", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Table 6 paths_summary", append=TRUE, row.names=FALSE)

>>>>

tar_load(Neale_SGG_dir_filt)
tar_load(traits_corr_spearman)
tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders_pos)
tar_load(corr_potential_trait_confounders_pos_MVMR_summary)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)
#tar_load(potential_trait_confounders_MVMR)
tar_load(corr_mat_traits)
tar_load(corr_impact_by_traits_causal)
tar_load(household_MR_SNPmeta_sensitivity)
tar_load(binned_pheno_corrs)

@@

sig_confounders <- corr_potential_trait_confounders_pos_MVMR_summary %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_MVMR) %>%
  mutate(total_potential_confounders = n()) %>%
  dplyr::summarize(description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding_MVMR)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  dplyr::rename(trait_description = description) %>% 
  dplyr::select(trait_description, trait_ID, everything())

sig_confounders_top5 <- corr_potential_trait_confounders_pos_MVMR_summary %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_MVMR) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 5) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(trait_description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   top5_potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding_MVMR)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  left_join(corr_vs_MR_effects_sig_dat.DT, by = c("trait_description", "trait_ID")) %>%
  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval,everything())

@@

write.table(corr_potential_trait_confounders_pos_MVMR_summary, "output/tables/C_summ_MVMR_result.csv")
# 
# corr_potential_trait_confounders_pos_MVMR_summary <- corr_potential_trait_confounders_pos_MVMR_summary %>% 
#   group_by(outcome_ID) %>%
#   mutate(corr_due_to_confounding_MVMR_summ = sum(corr_due_to_confounding_MVMR)) %>%
#   mutate(corr_due_to_confounding_summ = sum(corr_due_to_confounding))

@@

max_C_data <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% 
  #arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  #slice_head(n = 1) %>%
  left_join(couple_MR_vs_trait_corr %>% 
              dplyr::rename(outcome_ID_IVW_beta = IVW_beta) %>%
              dplyr::rename(outcome_ID_IVW_se = IVW_se) %>%
              dplyr::rename(outcome_ID_couple_r = couple_r) %>% 
              dplyr::rename(outcome_ID_couple_r_se = couple_r_se) %>% 
              dplyr::select(outcome_ID, outcome_ID_IVW_beta, outcome_ID_IVW_se, outcome_ID_couple_r, outcome_ID_couple_r_se)) %>%
  mutate(same_trait_corr_minus_MR_est = abs(outcome_ID_couple_r)- abs(outcome_ID_IVW_beta)) %>%
  dplyr::rename(Z_beta = IVW_beta) %>%
  dplyr::rename(Z_se = IVW_se) %>%
  dplyr::rename(Z_pval = IVW_pval) %>%
  mutate(Z_sq_beta = Z_beta^2) %>%
  mutate(Z_sq_se = sqrt(4*Z_beta^2*Z_se^2+2*Z_se^4)) %>%
  dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
  dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
  mutate(Z_prod = Z_sq_beta*Z_AM_beta) %>%
  mutate(Z_prod_var = variance_of_product(Z_sq_beta, Z_sq_se, Z_AM_beta, Z_AM_se)) %>%
  mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
  group_by(outcome_ID) %>% 
  mutate(corr_due_to_confounding_summ = sum(corr_due_to_confounding)) %>%
  mutate(corr_due_to_confounding_summ_SE = sqrt(sum(Z_prod_var))) %>%
  mutate(same_trait_corr_minus_MR_est_se = sqrt(variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))) %>%
  mutate(Phenotype = outcome_description) %>% 
  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
                               Phenotype == "Pulse rate, automated reading" ~ "Pulse rate",
                               Phenotype == "Body fat percentage" ~ "BFP",
                               Phenotype == "Standing height" ~ "Height",
                               Phenotype == "Reticulocyte percentage" ~ "Reticulocyte %",

@@

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - 1.96*corr_due_to_confounding_summ_SE, 
                 ymax = same_trait_corr_minus_MR_est + 1.96*corr_due_to_confounding_summ_SE),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = corr_due_to_confounding_summ - 1.96*Z_prod_se, 
                 xmax = corr_due_to_confounding_summ + 1.96*Z_prod_se),
    colour = "grey",
    width = 0
  ) +

@@

```{r large_results_table}

traits_description <- traits_final %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, r2, N_pairs, num_IVs_pass_het) %>% as_tibble() %>% 
  dplyr::rename(num_IVs = num_IVs_pass_het) %>%
  dplyr::rename(n_pairs = N_pairs) %>%
  dplyr::rename(pheno_couple_r2 = r2) %>% dplyr::rename(exposure_ID = Neale_pheno_ID)

# changed diff_p name 
MR_result_table <- corr_vs_MR_effects_dat %>% dplyr::select(exposure_ID, Correlation, couple_r_se, n_exposure, n_outcome, MR_est, IVW_se, IVW_pval, diff_p) %>%
  dplyr::rename(pheno_couple_r = Correlation) %>%
  dplyr::rename(pheno_couple_r_se = couple_r_se) %>%
  dplyr::rename(IVW_beta = MR_est) %>%
  dplyr::rename(couple_r_vs_IVW_diff_pval = diff_p)

AM_sex_spec_result <-
  household_MR_binned_het_sex_spec %>% filter(same_trait == TRUE) %>% dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()

AM_binned_result <- 
  household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% dplyr::select(exposure_ID, grouping_var, bin_slope_beta_wt, bin_slope_se_wt, bin_slope_pval_wt) %>% 
  mutate(grouping_var_rename = case_when(grouping_var=="age_even_bins" ~ "age",
                                         grouping_var=="time_together_even_bins" ~ "time_together")) %>% dplyr::select(-grouping_var) %>%
  dplyr::rename(bin_slope_beta_MR_wt = bin_slope_beta_wt) %>%
  dplyr::rename(bin_slope_se_MR_wt = bin_slope_se_wt) %>%
  dplyr::rename(bin_slope_pval_MR_wt = bin_slope_pval_wt)

sex_differences_summary <- AM_sex_spec_result %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval)) 
binned_summary <- AM_binned_result %>% pivot_wider(names_from = grouping_var_rename, values_from = c(bin_slope_beta_MR_wt, bin_slope_se_MR_wt, bin_slope_pval_MR_wt))

binned_corr_summary <- binned_pheno_corr_summary %>% 
  mutate(grouping_var_rename = case_when(group=="age_even_bins" ~ "age",
                                         group=="time_together_even_bins" ~ "time_together")) %>% unique() %>% dplyr::select(-group) %>%
  pivot_wider(names_from = grouping_var_rename, values_from = c(bin_slope_beta_pearson, bin_slope_se_pearson, 
                                                                bin_slope_pval_pearson,
                                                                bin_slope_beta_spearman, bin_slope_se_spearman, bin_slope_pval_spearman)) %>%
  dplyr::rename(exposure_ID = trait)

df_list <- list(traits_description, MR_result_table, sex_differences_summary, binned_summary, binned_corr_summary)
final_table <- df_list %>% reduce(full_join, by='exposure_ID') %>%
  dplyr::rename(trait_ID = exposure_ID) %>%
  dplyr::rename(trait_description = description) %>%
  dplyr::select(trait_description, trait_ID, everything())

@@

## Supplementary tables
write.xlsx_wrapper(final_table, file=supplementary_tables_file, sheetName="Supp Table 1 complete_summary", row.names=FALSE)
write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 2 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Supp Table 3 nominally_sig_sex_diff", row.names=FALSE)
write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 4 pheno_binned_corrs", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Supp Table 5 sig_corr_vs_MR", append = TRUE, row.names=FALSE)
write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Supp Table 6 sig_confounders", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Supp Table 7 paths_summary", append=TRUE, row.names=FALSE)
